Test Maps

### 2021-07-14 Solent Mapping_v2.Rmd
fixed yaml headers (now uses bookdown for floating ToC etc); fixed dataFolder for different platforms; checked maps plot what we think they do (some didn't); re-run

title: "Test Maps"
author: Meghan Kinglsey-Walsh
date: 'Last run at: 'r Sys.time()''
output:
output:
  pdf_document: default
  html_document: default
  word_document: default

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(sf)
library(spData)
library(spDataLarge)
library(dplyr)
dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/"
```

##Examine World Dataset
```{r Mapping population of the world (using ggplot2)}
head(world)
library(ggplot2)
ggplot2::ggplot(world)+
  geom_sf(aes(fill = pop))#draws blue map
ggplot2::ggplot(world) + 
  geom_sf(aes(fill = pop/1000000)) +
  scale_fill_continuous(name="Population (millions)",
                        low = "green",
                        high = "red") #changing colours

@@

##Mapping the Local Authority
###Mapping Domestic

####Mapping the Electricity Consumption
```{r loading in Local Authority electricity consumption data}
library(readr)
inFile <-paste0(dataFolder, "energy/Subnational_electricity_consumption_statistics_2019.csv")
la_elecData <- readr::read_csv(inFile)
summary(la_elecData)

```
```{r Loading in the Boundary Data for local authorities we are intrested in}
geo_query <- "https://ons-inspire.esriuk.com/arcgis/rest/services/Administrative_Boundaries/Local_Authority_Districts_December_2018_Boundaries_UK_BGC/MapServer/0/query?where=lad18nm%20IN%20(%27Southampton%27,%27Portsmouth%27,%27Winchester%27,%27Eastleigh%27,%27Isle%20of%20Wight%27,%27Fareham%27,%27Gosport%27,%27Test%20Valley%27,%27East%20Hampshire%27,%27Havant%27,%27New%20Forest%27,%27Hart%27,%27Basingstoke%20and%20Deane%27)&outFields=lad18cd,lad18nm,long,lat&outSR=4326&f=geojson"
#This only includes the ones we are interested in i.e the Solent
message("Loading LA geometry from ONS Open Geography API") #Specifying the location message
la_sf_data <- sf::st_read(geo_query)
head(la_sf_data)
```
```{r Mapping the local authority electricty consumption}
library(ggplot2)
la_elecData$lad18cd <- la_elecData$`LA Code`#Creating a variable with the local authority code in the same name of the consumption data
la_merged_sf <- merge(la_sf_data, la_elecData)#Merge the local Authority boundaries and the consumption data
ggplot2::ggplot(la_merged_sf) +
  geom_sf(aes(fill = Total_GWh))+scale_fill_gradient(low="green",high="red")

```
###Mapping Transport
```{r Loading the Transport consumption data}

@@

##Mapping the Middle Layer Super Output Areas (MSOA)
###Mapping Domestic
####Mapping the Electricity Consumption
```{r Loading in the MSOA electricity consumption data}
# electricity consumption data at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_DOM_ELEC_2019.csv")
msoa_elecData <- readr::read_csv(inFile)
head(msoa_elecData)
```

```{r Loading in the Boundary Data for the MSOA}
inf <-paste0(dataFolder, "boundaries/MSOA/msoa_solent.shp")
message("Loading Middle Layer Super Output Area (MSOA) from the ONS Open Georgraphy API")
#When using shp files you might also need to have other supplementary files in the same folder to allow it to run https://cfss.uchicago.edu/notes/simple-features/
msoa_sf_data <- sf::read_sf(inf) 
head (msoa_sf_data)
table(msoa_sf_data$LAD11NM)#how many MSOAs are in each LA
```
```{r Mapping the electrcity consumption of the MSOAs}
msoa_elecData$MSOA11CD <- msoa_elecData$ 'Middle Layer Super Output Area (MSOA) Code' #Creating a variable with the LA code and the name as in the sf_data
msoa_merged_sf <-merge(msoa_sf_data, msoa_elecData)
ggplot2::ggplot(msoa_merged_sf) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```

##Mapping the Lower Layer Super Output Area (LSOA)
###mapping Domestic
####Mapping Electricity consumption
```{r Loading in the LSOA electricty consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_ELEC_2019.csv")
lsoa_elecData <- readr::read_csv(inFile)
lsoa_elecData$LSOA11CD <- lsoa_elecData$`Lower Layer Super Output Area (LSOA) Code`
head(msoa_elecData)
```

```{r Loading in boundary data for the LSOA}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "boundaries/LSOA/lsoa_solent.shp")
message ("Loading LSOA boundaries from ONS Open geography API")
lsoa_sf_data <- sf::read_sf(inf)
head(lsoa_sf_data)
table(lsoa_sf_data$LAD11NM)#How many LSOAs are there in each LA
```
```{r Mapping the electricty consumption of the LSOAs}
lsoa_merged_sf <- merge(lsoa_sf_data, lsoa_elecData) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```
#####Looking at cities in more detail
```{r Looking at the cities in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
library(dplyr) # for filter, had to install this package
mapData <- dplyr::filter(lsoa_merged_sf, LAD11NM == "Southampton")

# plotting the elctricity consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Southampton")
```

###Mapping the Indices of Multiple Deprivation (JUST PLACE HOLDER DOESN'T WORK)
```{r Loading in the Indices of Multiple Deprivation data}
# Multiple deprivation data at MSOA pre downloaded
inFile <-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv")
lsoa_multidep <- readr::read_csv (inFile)
head(lsoa_multidep)
```
```{r Mapping the Indices of deprevation for the LSOA}
lsoa_multidep$LSOA11CD <- lsoa_multidep$lsoa11cd #creating a variable with the LA code in the same name as in sf_data
head(lsoa_multidep)
lsoa_multidep_merged_sf <- merge(lsoa_sf_data, lsoa_multidep) #merging these
ggplot2::ggplot(lsoa_multidep_merged_sf)+geom_sf(aes(fill=IMDScore))+scale_fill_continuous(name="(Text)IMDScore", low="green",high="red")+labs(caption ="Solent (all MSOAs)")
```

>>>>

---
title: "Test Maps"
author: "Meghan Kinglsey-Walsh"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # set this to FALSE if you don't want the code in the final oputput

# useful R libraries
library(ggplot2)
library(readr)
library(dplyr)

# GIS libraries
library(raster)
library(sf)
library(spData)
library(spDataLarge)

dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/" # default for Meghan
user <- Sys.info()[[7]] # who are we logged in as?
if(user == "ben"){
  dataFolder <- path.expand("~/University of Southampton/HCC Energy Landscape Mapping project - Documents/General/data/")
}

# this will not print anything out when knitted due to the include=FALSE setting in the chunk header
message("User: ", user)
message("dataFolder: ", dataFolder)
```

## Mapping the Local Authority

### Mapping Domestic

#### Mapping Electricity Consumption

```{r loading in Local Authority electricity consumption data}
inFile <-paste0(dataFolder, "energy/Subnational_electricity_consumption_statistics_2019.csv")
la_elecData <- readr::read_csv(inFile)
summary(la_elecData)
```

```{r Loading in the Boundary Data for local authorities we are intrested in}
# geo_query <- "https://ons-inspire.esriuk.com/arcgis/rest/services/Administrative_Boundaries/Local_Authority_Districts_December_2018_Boundaries_UK_BGC/MapServer/0/query?where=lad18nm%20IN%20(%27Southampton%27,%27Portsmouth%27,%27Winchester%27,%27Eastleigh%27,%27Isle%20of%20Wight%27,%27Fareham%27,%27Gosport%27,%27Test%20Valley%27,%27East%20Hampshire%27,%27Havant%27,%27New%20Forest%27,%27Hart%27,%27Basingstoke%20and%20Deane%27)&outFields=lad18cd,lad18nm,long,lat&outSR=4326&f=geojson"
# #This only includes the ones we are interested in i.e the Solent
# message("Loading LA geometry from ONS Open Geography API") #Specifying the location message
# la_sf_data <- sf::st_read(geo_query)

# The above is very clever but it takes time
# Load from file
la_shpFile <- paste0(dataFolder, "/boundaries/LA/la_solent.shp")
la_sf_data <- sf::st_read(la_shpFile)
head(la_sf_data)
```

```{r Mapping the local authority electricty consumption}
la_elecData$lad18cd <- la_elecData$`LA Code`#Creating a variable with the local authority code in the same name of the consumption data
la_merged_sf <- merge(la_sf_data, la_elecData)#Merge the local Authority boundaries and the consumption data
ggplot2::ggplot(la_merged_sf) +
  geom_sf(aes(fill = Total_GWh))+scale_fill_gradient(low="green",high="red")

@@

### Mapping Transport

#### Freight
```{r Loading the  Freight Transport consumption data for the Local Autority}
inFile <-paste0(dataFolder, "energy/Local Authority level Transport data.csv")
la_Transport <- readr::read_csv(inFile)
summary(la_Transport)
head(la_Transport)
```

```{r Mapping the local authority energy consumption (Freight)}

la_Transport$lad18cd <- la_Transport$Code # Creating a variable with the local authority code in the same name of the consumption data
la_merged_Transp <- merge(la_sf_data, la_Transport)# Merge the local Authority boundaries and the Transport data
ggplot2::ggplot(la_merged_Transp) +
  geom_sf(aes(fill = Freight)) +
  scale_fill_gradient(name = "Freight: Thousand tonnes of oil equivalent (ktoe)", low="green",high="red")
```

#### Personal

Note: this data is not per household/person so it will correlate (to some extent) with population size.

```{r Mapping the local authority energy consumption (Personal)}
# re-use the merged dataset
ggplot2::ggplot(la_merged_Transp) +
  geom_sf(aes(fill = Personal)) +
  scale_fill_gradient(name = "Personal: Thousand tonnes of oil equivalent (ktoe)", low="green",high="red") 
```

## Mapping the Middle Layer Super Output Areas (MSOA)

### Mapping Domestic

#### Mapping the Electricity Consumption

```{r Loading in the MSOA electricity consumption data}
# electricity consumption data at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_DOM_ELEC_2019.csv")
msoa_elecData <- readr::read_csv(inFile)
head(msoa_elecData)
```

```{r Loading in the Boundary Data for the MSOA}
inf <-paste0(dataFolder, "boundaries/MSOA/msoa_solent.shp")
message("Loading Middle Layer Super Output Area (MSOA) from file")
#When using shp files you might also need to have other supplementary files in the same folder to allow it to run https://cfss.uchicago.edu/notes/simple-features/
msoa_sf_data <- sf::read_sf(inf) 
head (msoa_sf_data)
table(msoa_sf_data$LAD11NM)#how many MSOAs are in each LA
```

```{r Mapping the electrcity consumption of the MSOAs}
msoa_elecData$MSOA11CD <- msoa_elecData$ 'Middle Layer Super Output Area (MSOA) Code' #Creating a variable with the LA code and the name as in the sf_data
msoa_merged_sf <-merge(msoa_sf_data, msoa_elecData)
ggplot2::ggplot(msoa_merged_sf) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Electricity: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```

#### Mapping the Gas
```{r Loading in the MSOA gas consumption data}
# gas consumption data at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_GAS_2019.csv")
msoa_gasData <- readr::read_csv(inFile)
head(msoa_gasData)
```
```{r Mapping the gas consumption of the MSOAs}
msoa_gasData$MSOA11CD <- msoa_gasData$ 'Middle Layer Super Output Area (MSOA) Code' 
msoa_merged_sf_domgas <-merge(msoa_sf_data, msoa_gasData)
ggplot2::ggplot(msoa_merged_sf_domgas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
> Does this match to non-gas areas??

### Mapping Commercial
#### Mapping Electricty
```{r Loading in the MSOA electricity consumption data for industry }
# electricity consumption data for industry at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_NONDOM_2019.csv")
msoa_elecDatanondom <- readr::read_csv(inFile)
head(msoa_elecDatanondom)
```

```{r Mapping the electrcity consumption of the MSOAs Industrial Sector}
msoa_elecDatanondom$MSOA11CD <- msoa_elecDatanondom$ 'Middle Layer Super Output Area (MSOA) Code' 
msoa_merged_sf_nondom <-merge(msoa_sf_data, msoa_elecDatanondom)
ggplot2::ggplot(msoa_merged_sf_nondom) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Electricity: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
#### Mapping Gas

```{r Loading in the MSOA gas consumption data for industry }
# gas consumption data for industry at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_NonDom_Gas_2019.csv")
msoa_gasDatanondom <- readr::read_csv(inFile)
head(msoa_gasDatanondom)
```

```{r Mapping the gas consumption of the MSOAs Industrial Sector}
msoa_gasDatanondom$MSOA11CD <- msoa_gasDatanondom$ 'Middle Layer Super Output Area (MSOA) Code' 
msoa_merged_sf_nondom_gas <-merge(msoa_sf_data, msoa_gasDatanondom)
ggplot2::ggplot(msoa_merged_sf_nondom_gas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
> The holes tell you where there is no non-domestic gas use

## Mapping the Lower Layer Super Output Area (LSOA)

### Mapping Domestic

#### Mapping Electricity consumption
```{r Loading in the LSOA electricty consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_ELEC_2019.csv")
lsoa_elecData <- readr::read_csv(inFile)
lsoa_elecData$LSOA11CD <- lsoa_elecData$`Lower Layer Super Output Area (LSOA) Code`
head(msoa_elecData)
```

```{r Loading in boundary data for the LSOA}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "boundaries/LSOA/lsoa_solent.shp")
message ("Loading LSOA boundaries from file")
lsoa_sf_data <- sf::read_sf(inf)
head(lsoa_sf_data)
table(lsoa_sf_data$LAD11NM)#How many LSOAs are there in each LA
```

```{r Mapping the electricty consumption of the LSOAs}
lsoa_merged_sf <- merge(lsoa_sf_data, lsoa_elecData) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```

##### Looking at cities in more detail
```{r Looking at the cities in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf, LAD11NM == "Southampton")

# plotting the electricity consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Southampton")
```

#### Mapping Gas

```{r Loading in the LSOA gas consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_GAS_2019.csv")
lsoa_gasData <- readr::read_csv(inFile)
lsoa_gasData$LSOA11CD <- lsoa_gasData$`Lower Layer Super Output Area (LSOA) Code`
head(lsoa_gasData)
```

```{r Mapping the gas consumption of the LSOAs}
lsoa_merged_sf_gas <- merge(lsoa_sf_data, lsoa_gasData) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf_gas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```

> Aha - this one shows where there is no gas :-)

### Mapping Industrial
#### Mapping Electricity consumption

> didn't we already do this in the MSOA section?

```{r Loading in the LSOA electricity consumption data for industry }
# electricity consumption data for industry at LSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_NONDOM_2019.csv")
msoa_elecDatanondom <- readr::read_csv(inFile)
head(msoa_elecDatanondom)
```

### Mapping the Indices of Multiple Deprivation

```{r Loading in the Indices of Multiple Deprivation data}
# Multiple deprivation data at MSOA pre downloaded
inFile <-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv")
lsoa_multidep <- readr::read_csv (inFile)
head(lsoa_multidep)
```

```{r Mapping the Indices of deprivation for the LSOA}
lsoa_multidep$LSOA11CD <- lsoa_multidep$`LSOA code (2011)` #creating a variable with the LSOA code in the same name as in sf_data
head(lsoa_multidep)
lsoa_multidep_merged_sf <- merge(lsoa_sf_data, lsoa_multidep) #merging these
ggplot2::ggplot(lsoa_multidep_merged_sf) +
  geom_sf(aes(fill=`Index of Multiple Deprivation (IMD) Rank`))+ # which data set has the scores?
  scale_fill_continuous(name="IMD Rank", low="green",high="red")+
  labs(caption ="Solent (all MSOAs)")
```

# The end

```{r environment}
R.Version()

@@
